Missing the Point\(\colon\) Non-Convergence in Iterative Imputation Procedures

### 2020-04-06 3.Thesis/1.SimulationStudy/Writeup/manuscript.Rmd
Feedback Gerko

---
title: Missing the Point$\colon$ Non-Convergence in Multiple Imputation Procedures
runninghead: Oberman
author:
- name: H. I. Oberman
  num: 1
address:
- num: 1
  org: Department of Methodology and Statistics, Utrecht University, Utrecht, The Netherlands
corrauth: "Hanne Oberman, Sjoerd Groenman building, Utrecht Science Park, Utrecht, The Netherlands."
email: h.i.oberman@uu.nl
abstract: "(**Rewrite:**) Multple imputation by chained equations (mice) is a widely used tool to accommodate missing data. While empirical evidence supports the validity of inferences obtained using mice, there is no consensus on the convergence properties of the method. This paper provides insight into non-convergence of mice algorithms."
keywords: MICE; convergence 
classoption:
  - Royal
  - times
bibliography: thesis.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
---

@@

At some point, any social scientist 
<!-- conducting statistical analyses  -->
will run into a missing data problem [@alli02]. Missingness is problematic because statistical inference cannot be performed on incomplete data without employing \emph{ad hoc} solutions (e.g., list-wise deletion), which may yield wildly invalid results [@buur18]. A popular method to obtain statistically valid results despite of missing data is 'multiple imputation' [MI; @rubin87]. This paper investigates a potential threat to the validity of this method: algorithmic non-convergence.

The multiple imputation method can be roughly divided into 3 steps. First, each missing value in an incomplete dataset is 'imputed' (i.e., filled in) several times. Next, the analysis of scientific interest is performed on each of the completed datasets. And finally, the results of these analyses are pooled into a single aggregated statistic. The theoretical foundation of these streps was laid by Rubin [-@rubin76, -@rubin87]. For an accessible and comprehensive introduction to MI from an applied perspective, see e.g. @buur18. In this paper, we focus on the first of these steps: the algorithmic procedure with which the missing values are imputed.

@@

There is no scientific consensus on the convergence properties of MI algorithms [@taka17]. Some default techniques might not yield converged states at all [@murr18]. Therefore, algorithmic convergence should be monitored carefully. The current practice is visually inspecting imputation chains for signs of non-convergence, but this may be challenging to the untrained eye [@buur18, $\S$ 6.5.2]. Moreover, visually inspecting imputation chains can only diagnose severely pathological cases of non-convergence. Therefore, diagnostic evaluation of convergence would be preferred.

Convergence cannot be determined for sure, because MICE (**introduce MICE before!**) is a type of MCMC method, and convergence of MCMC algorithms is not from a scalar to a point, but from one distribution to another. Therefore, generated values will vary even after convergence. The aim of convergence diagnostics for MCMC methods is thus not to establish the point at which convergence is reached, but to diagnose severe non-convergence.
<!-- Therefore we can only diagnose non-convergence, not convergence [@hoff09]. "a weak diagnostics is better than no diagnostic at all" [@cowl96]. -->

Several MCMC convergence diagnostics exist, but it is not known whetherthese are appropriate for MICE. The application of convergence diagnostics to MI methods has not been systematically studied [@buur18]. The open questions are, e.g.: How can non-convergence be diagnosed? How many iterations are sufficient/required? What are the effects of non-convergence on estimates, predictions and inference? What is an appropriate non-convergence diagnostic? Are the default number of iterations sufficient? The defaults of common MI software: SPSS = 10 [(link)](https://www.ibm.com/support/knowledgecenter/SSLVMB_24.0.0/spss/mva/syn_multiple_imputation_impute.html), Mplus = 100?? [(link)](https://pdfs.semanticscholar.org/e20e/29e008592cbfbaa567931f74cdfdb5451405.pdf?_ga=2.55354671.54033656.1584698748-527613517.1584698748), Stata = 10 [(link)](https://www.stata.com/manuals13/mi.pdf, p. 139), Amelia = NA, because resampling, not convergence [(link)](https://cran.r-project.org/web/packages/Amelia/Amelia.pdf), en MI = 30 [(link)](https://cran.r-project.org/web/packages/mi/mi.pdf).

@@

Let $y$ denote an $n \times p$ matrix containing the data values on $p$ variables for all $n$ units in a sample. The data value of unit $i$ ($i = 1, \dots, n$) on variable $j$ ($j = 1, \dots, p$) may be either observed or missing. The collection of observed data values in $y$ is denoted by $y_{obs}$; the missing part of $y$ is referred to as $y_{mis}$. (**This is not only notation, but theory too:** For each datapoint in $y_{mis}$ $M \times T$ times plausible values are sampled, where $M$ is the number of imputations ($m = 1, \dots, M$) and $T$ is the number of iterations ($t = 1, \dots, T$).) The collection of samples between the initial value (at $t=1$) and the imputed value (at $t=T$) will be referred to as an 'imputation chain'.

@@

Negative $AC$-values indicate divergence between subsequent samples. It may not be wise (**why? and how would we know? e.g., wider distrubutions? didn't seem like it...**) to terminating the algorithm when AC is negative. High $AC$-values are implausible in MI procedures. That is, the randomness induced by the MI algorithm effectively mitigates the risk of dependency within chains.

@@

and simulated with the function `mvtnorm::rmvnorm()`. Outcome variable $Y$ is subsequently calculated as 
$$Y =  1 + 2X_1 + .5X_2 - X_3 + \epsilon .$$
<!-- This note only considers a 'missing completely at random' (MCAR) mechanism, where the probability of being missing is equal for all $n \times p$ cells in $y$ [@rubin87].  -->

The complete data is 'amputed' once for each simulation repetition. That is, the function `mice::ampute()` is used to impose a missingness mechanism upon the data. (**change this:** The missingness is univariate, and the probability to be missing is the same for all four variables, namely 20\% (`prop = 0.8, mech = "MCAR"`). This leaves 20\% of the rows completely observed). **(Add different percentages missingness??)** This study only considers only an MCAR missingness mechanism. Therefore, results may not be extrapolated to other missing data problems. Proper performance of the convergence diagnostics under MCAR is necessary but not sufficient to demonstrate appropriateness of $\widehat{R}$ and $AC$ as convergence diagnostics. This is just a proof of concept.

Missing datapoints are imputed with the function `mice::mice()`. All MI procedures are performed with Bayesian linear regression imputation (`method = "norm"`), and five imputation chains (`m = 5`). The number of iterations varies between simulation conditions (`maxit = 1, 2, \dots, 100`).

@@

To estimate the quantity of scientific interest, $Q$, we perform multiple linear regression on each completed dataset with the function `stats::lm()`. We obtain an estimated regression coefficient per imputation, which are pooled into a single estimate, $\bar{Q}$. We use the function `mice::pool()` to get variance estimates according to Rubin's [-@rubin87] rules, and subsequently implement finite population pooling conform @vink14.

@@

## Univariate estimates and convergence diagnostics

>>>>

---
title: Missing the Point$\colon$ Non-Convergence in Iterative Imputation Procedures
runninghead: Oberman
author:
- name: H. I. Oberman
  num: 1
address:
- num: 1
  org: Department of Methodology and Statistics, Utrecht University, Utrecht, The Netherlands
corrauth: "Hanne Oberman, Sjoerd Groenman building, Utrecht Science Park, Utrecht, The Netherlands."
email: h.i.oberman@uu.nl
abstract: "(**Rewrite:**) Multple imputation by chained equations (mice) is a widely used tool to accommodate missing data. While empirical evidence supports the validity of inferences obtained using mice, there is no consensus on the convergence properties of the method. This paper provides insight into non-convergence of mice algorithms."
keywords: MICE; convergence 
classoption:
  - Royal
  - times
bibliography: thesis.bib
bibliographystyle: sageh
output:
  rticles::sage_article:
    keep_tex: yes
---

@@

Missing data problems are ubiquitous and pervasive. **or** Everybody who works with person data is likely to run into a missing data problem.  
<!-- At some point, any social scientist  -->
<!-- <!-- conducting statistical analyses  --> -->
<!-- will run into a missing data problem [@alli02].  -->
Missingness is problematic because statistical inference cannot be performed on the whole data without employing \emph{ad hoc} solutions (e.g., list-wise deletion), which may yield wildly invalid results [@buur18]. A popular method to obtain statistically valid results despite of missing data is 'multiple imputation' [MI; @rubin76]. The multiple imputation method can be roughly divided into 3 steps: First, each missing value in an incomplete dataset is 'imputed' (i.e., filled in) several times by some algorithmic procedure. Next, the analysis of scientific interest is performed on each of the completed datasets. And finally, the results of these analyses are pooled into a single aggregated statistic.

<!-- This paper investigates a potential threat to the validity of this method: algorithmic non-convergence.  -->

<!-- The theoretical foundation of these streps was laid by Rubin [-@rubin76, -@rubin87]. For an accessible and comprehensive introduction to MI from an applied perspective, see e.g. @buur18. In this paper, we focus on the first of these steps: the algorithmic procedure with which the missing values are imputed.  -->

Algorithms are often used to draw imputations from the posterior predictive distribution of the missing values. This introduces a potential threat to the validity of the imputations: what if the algorithm has not converged? Are the implications then to be trusted? And can we rely on the inference obtained on the completed data?

@@

There is no scientific consensus on the convergence properties of MI algorithms [@taka17]. Some default techniques (e.g., 'predictive mean modeling') might not yield converged states at all [@murr18]. Therefore, algorithmic convergence should be monitored carefully. The current practice is visually inspecting imputation chains for signs of non-convergence, but this may be challenging to the untrained eye [@buur18, $\S$ 6.5.2]. Moreover, visually inspecting imputation chains may only diagnose severely pathological cases of non-convergence. Therefore, diagnostic evaluation of convergence would be preferred.

Iterative imputation algorithms such as mice, are MCMC methods, and convergence of MCMC algorithms is not from a scalar to a point but from one distribution to another.  Therefore, generated values will vary even after convergence. The aim of convergence diagnostics for MCMC methods is thus not to establish the point at which convergence is reached, but to diagnose severe non-convergence.
<!-- Therefore we can only diagnose non-convergence, not convergence [@hoff09]. "a weak diagnostics is better than no diagnostic at all" [@cowl96]. -->

Several MCMC convergence diagnostics exist, but it is not known whetherthese are appropriate for MICE. The application of convergence diagnostics to MI methods has not been systematically studied [@buur18]. The open questions are, e.g.: How can non-convergence be diagnosed? How many iterations are sufficient/needed? What are the effects of non-convergence on estimates, predictions and inference? What is an appropriate non-convergence diagnostic? Are the default number of iterations sufficient? The defaults of common MI software: SPSS = 10 [(link)](https://www.ibm.com/support/knowledgecenter/SSLVMB_24.0.0/spss/mva/syn_multiple_imputation_impute.html), Mplus = 100?? [(link)](https://pdfs.semanticscholar.org/e20e/29e008592cbfbaa567931f74cdfdb5451405.pdf?_ga=2.55354671.54033656.1584698748-527613517.1584698748), Stata = 10 [(link)](https://www.stata.com/manuals13/mi.pdf, p. 139), Amelia = NA, because resampling, not convergence [(link)](https://cran.r-project.org/web/packages/Amelia/Amelia.pdf), en MI = 30 [(link)](https://cran.r-project.org/web/packages/mi/mi.pdf).

@@

Let $y$ denote an $n \times p$ matrix containing the data values on $p$ variables for all $n$ units in a sample. The data value of unit $i$ ($i = 1, \dots, n$) on variable $j$ ($j = 1, \dots, p$) may be either observed or missing. The collection of observed data values in $y$ is denoted by $y_{obs}$; the missing part of $y$ is referred to as $y_{mis}$. (**This is not only notation, but theory too:** For each datapoint in $y_{mis}$, we sample $M \times T$ times plausible values, where $M$ is the number of imputations ($m = 1, \dots, M$) and $T$ is the number of iterations ($t = 1, \dots, T$).) The collection of samples between the initial value (at $t=1$) and the final imputed value (at $t=T$) will be referred to as an 'imputation chain'.

@@

Negative $AC$-values indicate divergence within imputation chains. Subsequent sampled values within each imputation chain are less alike. It may not be wise (**why? and how would we know? e.g., wider distrubutions? didn't seem like it...**) to terminating the algorithm when AC is negative. **High $AC$-values are implausible in MI procedures. That is, the randomness induced by the MI algorithm effectively mitigates the risk of dependency within chains.**

@@

Outcome variable $Y$ is subsequently calculated as 
$$Y =  1 + 2X_1 + .5X_2 - X_3 + \epsilon .$$
<!-- This note only considers a 'missing completely at random' (MCAR) mechanism, where the probability of being missing is equal for all $n \times p$ cells in $y$ [@rubin87].  -->

**Add the function `mvtnorm::rmvnorm()` back in**

The complete data is 'amputed' once for each simulation repetition with function `mice::ampute(). (**change this:** The missingness is univariate, and the probability to be missing is the same for all four variables, namely 20\% (`prop = 0.8, mech = "MCAR"`). This leaves 20\% of the rows completely observed). **(Add different percentages missingness??)** This study only considers only an MCAR missingness mechanism. Therefore, results may not be extrapolated to other missing data problems. Proper performance of the convergence diagnostics under MCAR is necessary but not sufficient to demonstrate appropriateness of $\widehat{R}$ and $AC$ as convergence diagnostics. This is just a proof of concept.

Missing datapoints are imputed with the function `mice::mice()`. All MI procedures are performed with Bayesian linear regression imputation (`method = "norm"`), and five imputation chains (`m = 5`). The number of iterations varies between simulation conditions (`maxit = ` $$1, 2, \dots, 100$$).

@@

To estimate the quantity of scientific interest, $Q$, we perform multiple linear regression on each completed dataset with the function `stats::lm()`. We obtain an estimated regression coefficient per imputation, which are pooled into a single estimate, $\bar{Q}$. We use the function `mice::pool()` to estimate u bar (**check wat dat nou precies is!!**) according to Rubin's [-@rubin87] rules, and subsequently implement finite population pooling conform @vink14.

@@

## Univariate estimates and convergence diagnostics (theta = mean and theta = variance)

### 2020-04-06 3.Thesis/1.SimulationStudy/Writeup/manuscript.tex
Feedback Gerko

\title{Missing the Point\(\colon\) Non-Convergence in Multiple Imputation
Procedures}

@@

At some point, any social scientist will run into a missing data problem
\citep{alli02}. Missingness is problematic because statistical inference
cannot be performed on incomplete data without employing \emph{ad hoc}
solutions (e.g., list-wise deletion), which may yield wildly invalid
results \citep{buur18}. A popular method to obtain statistically valid
results despite of missing data is `multiple imputation'
\citep[MI;][]{rubin87}. This paper investigates a potential threat to
the validity of this method: algorithmic non-convergence.

The multiple imputation method can be roughly divided into 3 steps.
First, each missing value in an incomplete dataset is `imputed' (i.e.,
filled in) several times. Next, the analysis of scientific interest is
performed on each of the completed datasets. And finally, the results of
these analyses are pooled into a single aggregated statistic. The
theoretical foundation of these streps was laid by Rubin
\citeyearpar[\citeyearpar{rubin87}]{rubin76}. For an accessible and
comprehensive introduction to MI from an applied perspective, see
e.g.~\citet{buur18}. In this paper, we focus on the first of these
steps: the algorithmic procedure with which the missing values are
imputed.

There is no scientific consensus on the convergence properties of MI
algorithms \citep{taka17}. Some default techniques might not yield
converged states at all \citep{murr18}. Therefore, algorithmic
convergence should be monitored carefully. The current practice is
visually inspecting imputation chains for signs of non-convergence, but
this may be challenging to the untrained eye \citep[\(\S\)
6.5.2]{buur18}. Moreover, visually inspecting imputation chains can only
diagnose severely pathological cases of non-convergence. Therefore,
diagnostic evaluation of convergence would be preferred.

Convergence cannot be determined for sure, because MICE
(\textbf{introduce MICE before!}) is a type of MCMC method, and
convergence of MCMC algorithms is not from a scalar to a point, but from
one distribution to another. Therefore, generated values will vary even
after convergence. The aim of convergence diagnostics for MCMC methods
is thus not to establish the point at which convergence is reached, but
to diagnose severe non-convergence.

Several MCMC convergence diagnostics exist, but it is not known
whetherthese are appropriate for MICE. The application of convergence
diagnostics to MI methods has not been systematically studied
\citep{buur18}. The open questions are, e.g.: How can non-convergence be
diagnosed? How many iterations are sufficient/required? What are the
effects of non-convergence on estimates, predictions and inference? What
is an appropriate non-convergence diagnostic? Are the default number of
iterations sufficient? The defaults of common MI software: SPSS = 10
\href{https://www.ibm.com/support/knowledgecenter/SSLVMB_24.0.0/spss/mva/syn_multiple_imputation_impute.html}{(link)},
Mplus = 100??
\href{https://pdfs.semanticscholar.org/e20e/29e008592cbfbaa567931f74cdfdb5451405.pdf?_ga=2.55354671.54033656.1584698748-527613517.1584698748}{(link)},
Stata = 10
\href{https://www.stata.com/manuals13/mi.pdf,\%20p.\%20139}{(link)},
Amelia = NA, because resampling, not convergence
\href{https://cran.r-project.org/web/packages/Amelia/Amelia.pdf}{(link)},
en MI = 30
\href{https://cran.r-project.org/web/packages/mi/mi.pdf}{(link)}.

@@

Let \(y\) denote an \(n \times p\) matrix containing the data values on
\(p\) variables for all \(n\) units in a sample. The data value of unit
\(i\) (\(i = 1, \dots, n\)) on variable \(j\) (\(j = 1, \dots, p\)) may
be either observed or missing. The collection of observed data values in
\(y\) is denoted by \(y_{obs}\); the missing part of \(y\) is referred
to as \(y_{mis}\). (\textbf{This is not only notation, but theory too:}
For each datapoint in \(y_{mis}\) \(M \times T\) times plausible values
are sampled, where \(M\) is the number of imputations
(\(m = 1, \dots, M\)) and \(T\) is the number of iterations
(\(t = 1, \dots, T\)).) The collection of samples between the initial
value (at \(t=1\)) and the imputed value (at \(t=T\)) will be referred
to as an `imputation chain'.

@@

Negative \(AC\)-values indicate divergence between subsequent samples.
It may not be wise (\textbf{why? and how would we know? e.g., wider
distrubutions? didn't seem like it\ldots{}}) to terminating the
algorithm when AC is negative. High \(AC\)-values are implausible in MI
procedures. That is, the randomness induced by the MI algorithm
effectively mitigates the risk of dependency within chains.

@@

and simulated with the function \texttt{mvtnorm::rmvnorm()}. Outcome
variable \(Y\) is subsequently calculated as
\[Y =  1 + 2X_1 + .5X_2 - X_3 + \epsilon .\]

The complete data is `amputed' once for each simulation repetition. That
is, the function \texttt{mice::ampute()} is used to impose a missingness
mechanism upon the data. (\textbf{change this:} The missingness is
univariate, and the probability to be missing is the same for all four
variables, namely 20\% (\texttt{prop\ =\ 0.8,\ mech\ =\ "MCAR"}). This
leaves 20\% of the rows completely observed). \textbf{(Add different
percentages missingness??)} This study only considers only an MCAR
missingness mechanism. Therefore, results may not be extrapolated to
other missing data problems. Proper performance of the convergence
diagnostics under MCAR is necessary but not sufficient to demonstrate
appropriateness of \(\widehat{R}\) and \(AC\) as convergence
diagnostics. This is just a proof of concept.

Missing datapoints are imputed with the function \texttt{mice::mice()}.
All MI procedures are performed with Bayesian linear regression
imputation (\texttt{method\ =\ "norm"}), and five imputation chains
(\texttt{m\ =\ 5}). The number of iterations varies between simulation
conditions (\texttt{maxit\ =\ 1,\ 2,\ \textbackslash{}dots,\ 100}).

@@

To estimate the quantity of scientific interest, \(Q\), we perform
multiple linear regression on each completed dataset with the function
\texttt{stats::lm()}. We obtain an estimated regression coefficient per
imputation, which are pooled into a single estimate, \(\bar{Q}\). We use
the function \texttt{mice::pool()} to get variance estimates according
to Rubin's \citeyearpar{rubin87} rules, and subsequently implement
finite population pooling conform \citet{vink14}.

@@

\hypertarget{univariate-estimates-and-convergence-diagnostics}{%
\subsection{Univariate estimates and convergence
diagnostics}\label{univariate-estimates-and-convergence-diagnostics}}

>>>>

\title{Missing the Point\(\colon\) Non-Convergence in Iterative Imputation
Procedures}

@@

Missing data problems are ubiquitous and pervasive. \textbf{or}
Everybody who works with person data is likely to run into a missing
data problem.\\
--\textgreater{} Missingness is problematic because statistical
inference cannot be performed on the whole data without employing
\emph{ad hoc} solutions (e.g., list-wise deletion), which may yield
wildly invalid results \citep{buur18}. A popular method to obtain
statistically valid results despite of missing data is `multiple
imputation' \citep[MI;][]{rubin76}. The multiple imputation method can
be roughly divided into 3 steps: First, each missing value in an
incomplete dataset is `imputed' (i.e., filled in) several times by some
algorithmic procedure. Next, the analysis of scientific interest is
performed on each of the completed datasets. And finally, the results of
these analyses are pooled into a single aggregated statistic.

Algorithms are often used to draw imputations from the posterior
predictive distribution of the missing values. This introduces a
potential threat to the validity of the imputations: what if the
algorithm has not converged? Are the implications then to be trusted?
And can we rely on the inference obtained on the completed data?

There is no scientific consensus on the convergence properties of MI
algorithms \citep{taka17}. Some default techniques (e.g., `predictive
mean modeling') might not yield converged states at all \citep{murr18}.
Therefore, algorithmic convergence should be monitored carefully. The
current practice is visually inspecting imputation chains for signs of
non-convergence, but this may be challenging to the untrained eye
\citep[\(\S\) 6.5.2]{buur18}. Moreover, visually inspecting imputation
chains may only diagnose severely pathological cases of non-convergence.
Therefore, diagnostic evaluation of convergence would be preferred.

Iterative imputation algorithms such as mice, are MCMC methods, and
convergence of MCMC algorithms is not from a scalar to a point but from
one distribution to another. Therefore, generated values will vary even
after convergence. The aim of convergence diagnostics for MCMC methods
is thus not to establish the point at which convergence is reached, but
to diagnose severe non-convergence.

Several MCMC convergence diagnostics exist, but it is not known
whetherthese are appropriate for MICE. The application of convergence
diagnostics to MI methods has not been systematically studied
\citep{buur18}. The open questions are, e.g.: How can non-convergence be
diagnosed? How many iterations are sufficient/needed? What are the
effects of non-convergence on estimates, predictions and inference? What
is an appropriate non-convergence diagnostic? Are the default number of
iterations sufficient? The defaults of common MI software: SPSS = 10
\href{https://www.ibm.com/support/knowledgecenter/SSLVMB_24.0.0/spss/mva/syn_multiple_imputation_impute.html}{(link)},
Mplus = 100??
\href{https://pdfs.semanticscholar.org/e20e/29e008592cbfbaa567931f74cdfdb5451405.pdf?_ga=2.55354671.54033656.1584698748-527613517.1584698748}{(link)},
Stata = 10
\href{https://www.stata.com/manuals13/mi.pdf,\%20p.\%20139}{(link)},
Amelia = NA, because resampling, not convergence
\href{https://cran.r-project.org/web/packages/Amelia/Amelia.pdf}{(link)},
en MI = 30
\href{https://cran.r-project.org/web/packages/mi/mi.pdf}{(link)}.

@@

Let \(y\) denote an \(n \times p\) matrix containing the data values on
\(p\) variables for all \(n\) units in a sample. The data value of unit
\(i\) (\(i = 1, \dots, n\)) on variable \(j\) (\(j = 1, \dots, p\)) may
be either observed or missing. The collection of observed data values in
\(y\) is denoted by \(y_{obs}\); the missing part of \(y\) is referred
to as \(y_{mis}\). (\textbf{This is not only notation, but theory too:}
For each datapoint in \(y_{mis}\), we sample \(M \times T\) times
plausible values, where \(M\) is the number of imputations
(\(m = 1, \dots, M\)) and \(T\) is the number of iterations
(\(t = 1, \dots, T\)).) The collection of samples between the initial
value (at \(t=1\)) and the final imputed value (at \(t=T\)) will be
referred to as an `imputation chain'.

@@

Negative \(AC\)-values indicate divergence within imputation chains.
Subsequent sampled values within each imputation chain are less alike.
It may not be wise (\textbf{why? and how would we know? e.g., wider
distrubutions? didn't seem like it\ldots{}}) to terminating the
algorithm when AC is negative. \textbf{High \(AC\)-values are
implausible in MI procedures. That is, the randomness induced by the MI
algorithm effectively mitigates the risk of dependency within chains.}

@@

Outcome variable \(Y\) is subsequently calculated as
\[Y =  1 + 2X_1 + .5X_2 - X_3 + \epsilon .\]

\textbf{Add the function \texttt{mvtnorm::rmvnorm()} back in}

The complete data is `amputed' once for each simulation repetition with
function
\texttt{mice::ampute().\ (**change\ this:**\ The\ missingness\ is\ univariate,\ and\ the\ probability\ to\ be\ missing\ is\ the\ same\ for\ all\ four\ variables,\ namely\ 20\textbackslash{}\%\ (}prop
= 0.8, mech = ``MCAR''`). This leaves 20\% of the rows completely
observed). \textbf{(Add different percentages missingness??)} This study
only considers only an MCAR missingness mechanism. Therefore, results
may not be extrapolated to other missing data problems. Proper
performance of the convergence diagnostics under MCAR is necessary but
not sufficient to demonstrate appropriateness of \(\widehat{R}\) and
\(AC\) as convergence diagnostics. This is just a proof of concept.

Missing datapoints are imputed with the function \texttt{mice::mice()}.
All MI procedures are performed with Bayesian linear regression
imputation (\texttt{method\ =\ "norm"}), and five imputation chains
(\texttt{m\ =\ 5}). The number of iterations varies between simulation
conditions (\texttt{maxit\ =} \[1, 2, \dots, 100\]).

@@

To estimate the quantity of scientific interest, \(Q\), we perform
multiple linear regression on each completed dataset with the function
\texttt{stats::lm()}. We obtain an estimated regression coefficient per
imputation, which are pooled into a single estimate, \(\bar{Q}\). We use
the function \texttt{mice::pool()} to estimate u bar (\textbf{check wat
dat nou precies is!!}) according to Rubin's \citeyearpar{rubin87} rules,
and subsequently implement finite population pooling conform
\citet{vink14}.

@@

\hypertarget{univariate-estimates-and-convergence-diagnostics-theta-mean-and-theta-variance}{%
\subsection{Univariate estimates and convergence diagnostics (theta =
mean and theta =
variance)}\label{univariate-estimates-and-convergence-diagnostics-theta-mean-and-theta-variance}}
